---
title: "Description of loose.rock package"
author: "André Veríssimo"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
  github_document:
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
#
```

```{r gen.markdown, eval=FALSE, include=FALSE}
rmarkdown::render('./README.Rmd', output_format = 'github_document')
```

```{r, include=FALSE}
devtools::load_all(".")
```


# loose rock <img src="man/figures/loose.rock_logo.svg" width="120" align="right" />

> Set of useful functions in R that I reuse a lot

[![Travis-CI Build Status](https://travis-ci.org/averissimo/loose.rock.svg?branch=master)](https://travis-ci.org/averissimo/loose.rock)
[![Coverage status](https://codecov.io/gh/averissimo/loose.rock/branch/master/graph/badge.svg)](https://codecov.io/github/averissimo/loose.rock?branch=master)


### Install

The only pre-requirement is to install `biomaRt` bioconductor package as it cannot be installed automatically via CRAN.

All other dependencies should be installed when running the install command.

```{r install, eval=FALSE}
# install bioconductor
## try http:// if https:// URLs are not supported
source("https://bioconductor.org/biocLite.R")
biocLite()

# install the package
biocLite('averissimo/loose.rock', dependencies = TRUE)

# use the package
library(loose.rock)
```

### Overview

With personal functions I like to reuse everytime!

- `draw.kaplan()` : Draw kaplan curves based on Prognostic Index of Risk (calculated by coxph or something else)
- `coding.genes()`: downloads protein coding genes from external databases
- `gen.synth.xdata()`: generate random matrix with pre-determined covariance
- `balanced.cv.folds()` and `balanced.train.and.test()`: get balanced train/test sets and cv folds.
- `run.cache()`: keep cache or results of a function
- `proper()` : Capitalize string using regexpression
- `my.colors()` : My own pallete
- `my.symbols()` : Same with symbols to plots
- ... check out rest of R folder

## draw.kaplan

```{r, message=FALSE, warning=FALSE}
library(survival)
library(dplyr)
library(ggfortify)
library(gridExtra)
```

```{r draw.kaplan}
data(flchain)
ydata <- data.frame( time = flchain$futime, status = flchain$death)
xdata <- cbind(flchain$age, as.numeric(flchain$sex == 'M') * 2 + 1, flchain$sample.yr, flchain$kappa)
page <- draw.kaplan(list(Age= c(1,0,0,0)), xdata = xdata, ydata = ydata)$plot
psex <- draw.kaplan(list(Sex= c(0,1,0,0)), xdata = xdata, ydata = ydata)$plot
grid.arrange(page, psex, ncol = 2)
#
draw.kaplan(list(Age= c(1,0,0,0), Sex = c(0,1,0,0)), xdata = xdata, ydata = ydata)$plot
```

## Get a current list of protein coding genes

```{r coding.genes, collapse=TRUE}
genes <- coding.genes()
genes %>%
  arrange(external_gene_name) %>%
  head(n = 30) %>%
  knitr::kable()
```


## Balanced test/train dataset

This is specially relevant in survival or binary output with few cases of one category that need to be well distributed among test/train datasets or in cross-validation folds.

Example below sets aside 90% of the data to the training set. As samples are already divided in two sets (`set1` and `set2`), it performs the 90% separation for each and then joins (with option `join.all = T`) the result.

```{r balanced.sets, results='hold'}
set1 <- c(T,T,T,T,T,T,T,T,F,T,T,T,T,T,T,T,T,T,F,T)
set2 <- !set1
cat('Set1\n', set1, '\n\nSet2\n', set2, '\n\nTraining / Test set using logical indices\n\n')
set.seed(1985)
balanced.train.and.test(set1, set2, train.perc = .9)
#
set1 <- which(set1)
set2 <- which(set2)
cat('##### Same sets but using numeric indices\n\n', 'Set1\n', set1, '\n\nSet2\n', set2, '\n\nTraining / Test set using numeric indices\n')
set.seed(1985)
balanced.train.and.test(set1, set2, train.perc = .9)
#
```

## Generate synthetic matrix with covariance

```{r gen.synth}
xdata1 <- gen.synth.xdata(10, 5, .2)
xdata2 <- gen.synth.xdata(10, 5, .75)
```

```{r show.gen.synth, echo=FALSE}
#
cat('Using .2^|i-j| to generate co-variance matrix\n\n')
cat('X generated\n\n')
data.frame(xdata1)
cat('cov(X)\n\n')
data.frame(cov(xdata1))
draw.cov.matrix(xdata1) + ggplot2::ggtitle('X1 Covariance Matrix')
#
cat('Using .75^|i-j| to generate co-variance matrix (plotting correlation)\n\n')
cat('X generated\n\n')
data.frame(xdata2)
cat('cov(X)\n\n')
data.frame(cor(xdata2, method = 'pearson'))
draw.cov.matrix(xdata2, fun = cor, method = 'pearson') + ggplot2::ggtitle('X2 Pearson Correlation Matrix')
```

## Save in cache

Uses a cache to save and retrieve results. The cache is automatically created with the arguments and source code for function, so that if any of those changes, the cache is regenerated.

Caution: Files are not deleted so the cache directory can become rather big.

```{r runcache1}
a <- run.cache(sum, 1, 2)
b <- run.cache(sum, 1, 2)
all(a == b)
```

```{r runcache2}
a <- run.cache(rnorm, 5, seed = 1985)
b <- run.cache(rnorm, 5, seed = 2000)
all(a == b)
```

run-cache was originaly intended to be used to calculate big correlation matrix

```{r runcache-benchmark}
n.rows <- 1000
n.cols <- 50000
xdata <- matrix(rnorm(n.rows * n.cols), ncol = n.cols)
# making sure cache is saved
.Last.value <- run.cache(sapply, 2:n.cols, function(ix) {cor(xdata[,1], xdata[,ix])})
run.cache.digest <- list(digest.cache(xdata))
my.fun <- function(ix) {cor(xdata[,1], xdata[,ix])}
microbenchmark::microbenchmark(
  run.cche.non.cached    = run.cache(sapply, 2:n.cols, my.fun, show.message = FALSE, force.recalc = T),
  run.cache.cached       = run.cache(sapply, 2:n.cols, my.fun, show.message = FALSE),
  run.cache.cached.speed = run.cache(sapply, 2:n.cols, my.fun, cache.digest = run.cache.digest, show.message = FALSE),
  actual.function        = sapply(2:n.cols, my.fun), 
  actual.4cores          = unlist(parallel::mclapply(2:n.cols, my.fun, mc.cores = 4)),
  times = 5)
```

## Proper

One of such is a proper function that capitalizes a string.

```{r proper}
x <- "OnE oF sUcH iS a proPer function that capitalizes a string."
proper(x)
```

## Custom colors and symbols

`my.colors()` and `my.symbols()` can be used to improve plot readability.

```{r mycolors}
xdata <- -10:10
plot(xdata, 1/10 * xdata * xdata + 1, type="l", pch = my.symbols(1), col = my.colors(1), cex = .9,
     xlab = '', ylab = '', ylim = c(0, 20))
grid(NULL, NULL, lwd = 2) # grid only in y-direction
for (ix in 2:22) {
  points(xdata, 1/10 * xdata * xdata + ix, pch = my.symbols(ix), col = my.colors(ix), cex = .9)
}
```
