---
title: "Overview of loose.rock"
author: "André Veríssimo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Overview of loose.rock}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, eval=FALSE, include=FALSE}
# before running this, change ../man/figures/loose.rock_logo.svg" width="120" align="right
# and uncomment fig.path = "man/figures/README-"
rmarkdown::render('Overview.Rmd', output_file = file.path(dirname(getwd()), 'README.md'), output_format = 'github_document')
```


<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(
  # fig.path = "man/figures/README-",
  collapse = TRUE,
  comment = "#>"
)
#
```

# loose rock 
<img src="../man/figures/loose.rock_logo.svg" width="120" align="right" />

> Set of Functions to Use in Survival Analysis and in Data Science

![R CMD check](https://github.com/averissimo/loose.rock/workflows/R-CMD-check/badge.svg)
[![Coverage Status](https://codecov.io/gh/averissimo/loose.rock/branch/master/graph/badge.svg)](https://codecov.io/github/averissimo/loose.rock?branch=master)
[![CRAN Version](https://img.shields.io/cran/v/loose.rock.svg)](https://cran.r-project.org/package=loose.rock)


```{r, include=FALSE}
library(loose.rock)
```

Collection of function to improve workflow in survival analysis and data 
science. Among the many features, the generation of balanced datasets, 
retrieval of protein coding genes from two public databases (live) and 
generation of random matrix based on covariance matrix.

The work has been mainly supported by two grants: FCT SFRH/BD/97415/2013 
and the EU Commission under SOUND project with contract number 633974.

### Install

The only pre-requirement is to install `biomaRt` bioconductor package as 
it cannot be installed automatically via CRAN.

All other dependencies should be installed when running the install command.

```{r install, eval=FALSE}
if (!require("BiocManager"))
  install.packages("BiocManager")
BiocManager::install("loose.rock")

# use the package
library(loose.rock)
```

### Overview

- `coding.genes()`: downloads protein coding genes from external databases
- `gen.synth.xdata()`: generate random matrix with pre-determined covariance
- `balanced.cv.folds()` and `balanced.train.and.test()`: get balanced 
train/test sets and cv folds.
- `run.cache()`: keep cache or results of a function
- `proper()` : Capitalize string using regexpression
- `my.colors()` : My own pallete
- `my.symbols()` : Same with symbols to plots
- ... check out rest of Documentation

## Libraries required for this vignette

```{r, message=FALSE, warning=FALSE}
library(dplyr)
```

## Get a current list of protein coding genes

Showing only a random sample of 15

```{r coding.genes, collapse=TRUE, message=FALSE, warning=FALSE}
coding.genes() %>%
  arrange(external_gene_name) %>% {
   slice(., sample(seq(nrow(.)), 15)) 
  } %>%
  knitr::kable()
```

## Balanced test/train dataset

This is specially relevant in survival or binary output with few cases of 
one category that need to be well distributed among test/train datasets or 
in cross-validation folds.

Example below sets aside 90% of the data to the training set. As samples are 
already divided in two sets (`set1` and `set2`), it performs the 90% separation 
for each and then joins (with option `join.all = T`) the result.

```{r balanced.sets, results='hold'}
set1 <- c(rep(TRUE, 8), FALSE, rep(TRUE, 9), FALSE, TRUE)
set2 <- !set1
cat(
  'Set1', '\n', set1, '\n\n',
  'Set2', '\n', set2, '\n\n',
  'Training / Test set using logical indices', '\n\n'
)
set.seed(1985)
balanced.train.and.test(set1, set2, train.perc = .9)
#
set1 <- which(set1)
set2 <- which(set2)
cat(
  '##### Same sets but using numeric indices', '\n\n', 
  'Set1', '\n', set1, '\n\n', 
  'Set2', '\n', set2, '\n\n', 
  'Training / Test set using numeric indices', '\n')
set.seed(1985)
balanced.train.and.test(set1, set2, train.perc = .9)
#
```

## Generate synthetic matrix with covariance

```{r gen.synth}
xdata1 <- gen.synth.xdata(10, 5, .2)
xdata2 <- gen.synth.xdata(10, 5, .75)
```

```{r show.gen.synth, echo=FALSE}
#
cat('Using .2^|i-j| to generate co-variance matrix\n\n')
cat('X generated\n\n')
data.frame(xdata1)
cat('cov(X)\n\n')
data.frame(cov(xdata1))
draw.cov.matrix(xdata1) + ggplot2::ggtitle('X1 Covariance Matrix')
#
cat('Using .75^|i-j| to generate co-variance matrix (plotting correlation)\n\n')
cat('X generated\n\n')
data.frame(xdata2)
cat('cov(X)\n\n')
data.frame(cor(xdata2, method = 'pearson'))
draw.cov.matrix(xdata2, fun = cor, method = 'pearson') + 
  ggplot2::ggtitle('X2 Pearson Correlation Matrix')
```

## Save in cache

Uses a cache to save and retrieve results. The cache is automatically created 
with the arguments and source code for function, so that if any of those 
changes, the cache is regenerated.

*Caution: Files are not deleted so the cache directory can become rather big.*

Set a temporary directory to save all caches (optional)

```{r}
base.dir(tempdir())
```

Run sum function twice

```{r runcache1}
a <- run.cache(sum, 1, 2)
b <- run.cache(sum, 1, 2)
all(a == b)
```

Run rnorm function with an explicit seed 
*(otherwise it would return the same random number)*

```{r runcache2}
a <- run.cache(rnorm, 5, seed = 1985)
b <- run.cache(rnorm, 5, seed = 2000)
all(a == b)
```

## Proper

One of such is a proper function that capitalizes a string.

```{r proper}
x <- "OnE oF sUcH iS a proPer function that capitalizes a string."
proper(x)
```

## Custom colors and symbols

`my.colors()` and `my.symbols()` can be used to improve plot readability.

```{r mycolors}
xdata <- -10:10
plot(
  xdata, 1/10 * xdata * xdata + 1, type="l", 
  pch = my.symbols(1), col = my.colors(1), cex = .9,
  xlab = '', ylab = '', ylim = c(0, 20)
)
grid(NULL, NULL, lwd = 2) # grid only in y-direction
for (ix in 2:22) {
  points(
    xdata, 1/10 * xdata * xdata + ix, pch = my.symbols(ix), 
    col = my.colors(ix), cex = .9
  )
}
```
